Comment: Inferring Trade Costs from Trade Booms and Trade Busts

### 2017-09-26 Paper/Manuscript.tex
Update Manuscript.tex

This paper starts by presenting Jacks \textit{et al.}'s approach of the
measure of trade costs, and insists on its relevance at the
bilateral level to control for multilateral trade barriers in
gravity regressions (section \ref{1}). It then highlights that the
result on a difference of nature between the two globalizations
is paradoxical since it cannot be deduced from a comparison of the
evolution of openness ratios. Section \ref{2} shows that the
conclusion is only driven by the authors' \textit{ad hoc} aggregation
method. We propose a microfounded way to aggregate trade costs
and the puzzle fades away. Section \ref{3} explores the reasons
why Jacks \textit{et al.}'s aggregation technique ends up providing
different results for the two globalizations. We argue that
what Jacks \textit{et al.} misleadingly attribute to unequal trade costs
decreases between the two globalizations is instead a
difference in the distribution of trade growth over trading
partners.

@@

Where $s_i=y_i/(y_i+y_j)$.

@@

Examining this equation, it is clear that the method consists in determining the evolution of one unobservable variable (trade costs) based on the evolution of two groups of observable variables (trade flows and output) and a number of hypothesis on preferences and parameters ($\sigma$ and the relationship between $x_{ii}$, output and trade). If one assumes that preferences and parameters do not change (though these points are partly addressed  in appendix B of  \cite{JMN2011}), all changes in trade costs must restate information contained in changes in the measure of trade flows relative to economic size, like an openness ratio defined for instance as the ratio of exportations to GDP.

@@

\caption{Decomposition of the growth in international trade
(logarithms) with \textit{ad hoc} averages and a microfounded
aggregation method. First wave of globalization,
1870-1913. }
\Note {\emph{JMN 2011} refers to the averaging over dyads,
\emph{JMN by country} by country refers to the averaging over
trading partners for one country, \emph{our methodl}
refers to the aggregation method we offer.

@@

\end{tabularx}
\caption{Decomposition of the growth in international trade
(logarithms) with \textit{ad hoc} averages and a microfounded
aggregation method. Second wave of globalization, 1950-2000.}
\Note {\emph{JMN 2011} refers to the averaging over dyads,
\emph{JMN by country} by country refers to the averaging over
trading partners for one country, \emph{our methodl}
refers to the aggregation method we offer.

@@

The reason why Jacks \textit{et al.}'s aggregate measure of trade costs yields different conclusion the micro-founded method (and openness rate) is its sensitivity to structure
effects. To explain this idea, and for clarity purposes, let us move to a world of symmetric partners (domestic trade is equal in $i$ and $j$ and $x_{ij}$=$x_{ji}$).  Dyadic bilateral trade flows are not equal because some dyads are further appart than others. Equation (\ref{keyeq}) gives:

@@

Note $\overline{a_j}$ the arithmetic mean
of $a_j$ over $j$ ($\overline{a_j}=1/n \sum_{j=1}^{n}a_j$). One can
then compare the unweighted average of $\tau_{ij}$ according to equation 
(\ref{tausimplified}) and the measure derived from the
two-countries model in equation (\ref{keyeq2model}):

@@

Let us now move away from our symmetric world. Jacks  \textit{et al.} partially correct that bias by using a end-of-period-GDP weights average to compute the average of  $\tau_{ij}$.
However, as tables \ref{glob1} and \ref{glob2} show, weighting by the GDP of trade partners, besides not being theoretically justified, does not provide an accurate correction of the bias in the measure.
One reason for this failure is that the relationship between GDP and the importance of trade flows is not systematic.

For these reasons, Jacks  \textit{et al.}'s average of the bilateral trade costs $\tau_{ij}$ captures both the evolution of trade costs and the distribution of the evolution of trade costs over dyad sizes.
As a result, if dyads with small starting trade experience faster growth of trade than others, the decline of trade costs measured by $\tau_i$ will be overestimated compared to the decline of $\upsilon_i$.
This fits the data and can explain the difference they find between the first and the second globalization. In the first globalization, pairs of countries that initially traded little together experienced relative faster trade growth.
For both globalizations, figure \ref{structure} plots the growth of trade flows (measured by $\Delta ln (\sqrt{x_{ij} x_{ji}})$) as a function of the initial value of trade (measured by the logarithm of the geometric average of bilateral average $ln(\sqrt{x_{ij} x_{ji}})$) for all dyads in the sample.

>>>>

In this paper, we first present Jacks \textit{et al.}'s approach of the
measure of trade costs, and insists on its relevance at the
bilateral level to control for multilateral trade barriers in
gravity regressions (section \ref{1}). We then highlight that the
result on the difference of nature between the two globalizations
is paradoxical since it cannot be deduced from a comparison of the
evolution of openness ratios. Section \ref{2} shows that the
conclusion is only driven by the authors' \textit{ad hoc} aggregation
method. We propose a microfounded way to aggregate trade costs
and the puzzle fades away. Section \ref{3} explores the reasons
why Jacks \textit{et al.}'s aggregation technique ends up providing
different results for the two globalizations. We argue that
what Jacks \textit{et al.} misleadingly attribute to unequal trade costs
decreases between the two globalizations is instead a
difference in the distribution of trade growth over trading
partners.

@@

where $s_i=y_i/(y_i+y_j)$.\\

@@

Examining this equation, it is clear that the method consists in determining the evolution of one unobservable variable (trade costs) based on the evolution of two groups of observable variables (trade flows and output) and a number of hypotheses on preferences and parameters ($\sigma$ and the relationship between $x_{ii}$, output and trade). If one assumes that preferences and parameters do not change (though these points are partly addressed  in appendix B of  \cite{JMN2011}), all changes in trade costs must restate information contained in changes in the measure of trade flows relative to economic size, like an openness ratio defined for instance as the ratio of exportations to GDP.

@@

\caption{Decomposition of the growth in international trade
(logarithms) with \textit{ad hoc} averages and a microfounded
aggregation method. First wave of globalization,
1870-1913. }
\Note {\emph{JMN 2011} refers to the averaging over dyads,
\emph{JMN by country} by country refers to the averaging over
trading partners for one country, \emph{our method}
refers to the aggregation method we offer.

@@

\end{tabularx}
\caption{Decomposition of the growth in international trade
(logarithms) with \textit{ad hoc} averages and a microfounded
aggregation method. Second wave of globalization, 1950-2000.}
\Note {\emph{JMN 2011} refers to the averaging over dyads,
\emph{JMN by country} by country refers to the averaging over
trading partners for one country, \emph{our method}
refers to the aggregation method we offer.

@@

The reason why Jacks \textit{et al.}'s aggregate measure of trade costs yields different conclusions compared to the micro-founded method (and openness rate) is its sensitivity to structure
effects. To explain this idea, and for clarity purposes, let us move to a world of symmetric partners (domestic trade is equal in $i$ and $j$ and $x_{ij}$=$x_{ji}$).  Dyadic bilateral trade flows are not equal because some dyads are further apart than others. Equation (\ref{keyeq}) gives:

@@

Note $\overline{a_j}$ the arithmetic mean
of $a_j$ over $j$ ($\overline{a_j}=1/n \sum_{j=1}^{n}a_j$). One can
then compare the unweighted average of $\tau_{ij}$ according to equation 
(\ref{tausimplified}) and the measure derived from the
two-country model in equation (\ref{keyeq2model}):

@@

Let us now move away from our symmetric world. Jacks  \textit{et al.} partially correct that bias by using a end-of-period-GDP weighted average to compute the average of  $\tau_{ij}$.
However, as tables \ref{glob1} and \ref{glob2} show, weighting by the GDP of trade partners, besides not being theoretically justified, does not provide an accurate correction of the bias in the measure.
One reason for this failure is that the relationship between GDP and the importance of trade flows is not systematic.

For these reasons, Jacks  \textit{et al.}'s average of the bilateral trade costs $\tau_{ij}$ captures both the evolution of trade costs and the distribution of the evolution of trade costs over dyad sizes.
As a result, if dyads with small starting trade experience faster growth of trade than others, the decline of trade costs measured by $\tau_i$ will be overestimated compared to the decline of $\upsilon_i$.
This fits the data and can explain the difference they find between the first and the second globalization. For both globalizations, figure \ref{structure} plots the growth of trade flows (measured by $\Delta ln (\sqrt{x_{ij} x_{ji}})$) as a function of the initial value of trade (measured by the logarithm of the geometric average of bilateral average $ln(\sqrt{x_{ij} x_{ji}})$) for all dyads in the sample.\\
